Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2024-12-20 analysis/paper/article.qmd
stash

The aim of this paper is to TEACH JESS HOW TO USE QUARTO AND GITHUB, UNFORTUNATELY FOR JESS. We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.

@@

Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from. Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing]. Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019]. This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology. This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.

>>>>

The aim of this paper is to explore the demand-side of the academic job market for archaeologists in the United States. We had two aims: to determine if disciplinary trends can be discerned in the topical, geographic, and method foci of the positions advertised; and to investigate how the requirements for applicants have changed over time.

@@

Studies of who holds tenure-track faculty positions in archaeology at US universities, shows that success in job applications is strongly determined by where applicants get their PhD from. <!-- here is a comment in the middle --> Data from the 2014–2015 AnthroGuide publication of the American Anthropological Association shows that just ten out of over 100 US graduate programs produced over 30% of the graduates hired into tenure-track faculty positions [@speakman2018choosing]. Similarly, a network analysis of 1,918 faculty holding tenured or tenure-track positions at PhD-granting anthropology programs in the US in 2015 (including 506 archaeologists) found that the just fifteen graduate programs produced 52.8 percent of tenured and tenure-track positions [@kawaSocialNetworkUS2019]. This network analysis further found that programs with large endowments and with faculty who hold prestigious awards and are widely cited by other scholars produce the majority of tenured and tenure-track faculty in anthropology. This pattern of inequality is not unique to archaeology [@wapmanQuantifyingHierarchyDynamics2022a], but these studies draw attention to the paradox of a discipline that has many members committed to fighting social inequalities, and yet continues to reproduce its own systematic inequalities and hierarchies through hiring practices that favour an elite minority of graduates with prestigious institutional affiliations.
